---
title: "Analyzing NOAA's storm data"
author: "Me"
date: "12/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Title: Analyzing NOAA's storm data

# Synopsis

This assignment aims to explore and manipulate the  U.S. National Oceanic and Atmospheric Administration's (NOAA) storm data to answer the following questions:

* Across the United States, which types of events are most harmful with respect to population health?
* Across the United States, which types of events have the greatest economic consequences?

### Summary

We used the NOAA's storm data to identify the most damaging events to both human life and economic properties. We have identified that the storm is the deadliest event recording to a total of around 100,000 injuries and lives taken. For economic damage, flood has the highest property damage while drought has the highest crop damage


# Data Exploration and Processing

### Load data and libraries needed

Load the libraries needed for data analysis
```{r}
library(tidyverse)
library(janitor)
library(ggplot2)
library(patchwork)
```


Download the data from the given URL and upload it to R
```{r}
file_url = 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2'
download.file(file_url, destfile = paste0(getwd(), '/data.csv.bz2'))
data = data.table::fread('data.csv.bz2')
```


### Exploring the data

Check the values on each column
```{r}
head(data[,1:5])
```

```{r}
dim(data)
```
There are more than 900k rows and 37 columns.


Check the columns
```{r}
colnames(data)
```


### Select appropriate columns

Select only those columns that are needed
```{r}
cols_needed = c('EVTYPE', 'FATALITIES', 'INJURIES', "PROPDMG", 'PROPDMGEXP','CROPDMG', 'CROPDMGEXP')

data = data[, ..cols_needed]
data = clean_names(data)
```
Check structure of selected columns

```{r}
str(data)
```


### Format exponent data

Let's format first hte property damage exponent and crop damage exponent

There's a need to make property damage exponent and crop damage exponent fixed

Make all upper case for property damage exponent and crop damage exponent
```{r}
data = data %>%
      mutate(propdmgexp = toupper(propdmgexp),
             cropdmgexp = toupper(cropdmgexp))
```

Convert letters to numbers
```{r}
data = data %>%
     mutate(propdmgexp = case_when(propdmgexp == 'H' ~ '2',
                                   propdmgexp == 'K' ~ '3',
                                   propdmgexp == 'M' ~ '6',
                                   propdmgexp == 'B' ~ '9',
                                   TRUE ~ propdmgexp),
            cropdmgexp = case_when(cropdmgexp == 'H' ~ '2',
                                   cropdmgexp == 'K' ~ '3',
                                   cropdmgexp == 'M' ~ '6',
                                   cropdmgexp == 'B' ~ '9',
                                   TRUE ~ cropdmgexp))
```



Convert to numeric and fill others with zero
```{r}
data = data %>%
      mutate(propdmgexp = parse_number(propdmgexp),
             cropdmgexp = parse_number(cropdmgexp))

data[is.na(data$propdmgexp),'propdmgexp'] = 0
data[is.na(data$cropdmgexp),'cropdmgexp'] = 0
```

```{r}

data %>%
      mutate(propdmg1 = propdmg * (10^propdmgexp), 
             cropdmg1 = cropdmg * (10^ cropdmgexp))  %>%
      arrange(desc(cropdmg))
```

### Calculate actual cost

Calculate the actual property damage and crop damage
```{r}
data = data %>%
      mutate(propdmg = propdmg * (10^propdmgexp), 
             cropdmg = cropdmg * (10^cropdmgexp)) %>%
      select(-propdmgexp, -cropdmgexp)
```




### Create dataframe for total health and total economic damages

Create a column for health damage and economic damage by combininig number of fatalities and  injuries and property damage and crop damage
```{r}
health_economic = data %>%
      mutate(health_damage = fatalities + injuries,
             economic_damage = propdmg + cropdmg) %>%
      select(evtype, health_damage, economic_damage)
```

### Create a dataframe for separate property damage and crop damage

Combine the economic damage and create an identifier
```{r}
economic_damages = data %>%
      gather(key = 'type', value = 'damage', propdmg, cropdmg) %>%
      select(evtype, type, damage)
```


# Results

### Question 1: Most frequent events

```{r}
graph1 = data %>%
      mutate(evtype = fct_lump(evtype, 15)) %>%
      count(evtype) %>%
      mutate(evtype = fct_reorder(evtype, n)) %>%
      ggplot(aes(evtype, n))+
      geom_col(aes(fill = evtype)) +
      coord_flip()+
      labs(title = 'Occurences of natural events in the US',
           subtitle = 'Hail is the most occuring natural event',
           x = '',
           y = 'count')+
      scale_y_continuous(labels = scales::comma)+
      theme_minimal()+
      theme(legend.position = 'none')

graph1
```
This first image presents the counts of natural events as recorded by U.S. National Oceanic and Atmospheric Administration's (NOAA)

### Question2:  Deadliest events

```{r}
health_economic_total = health_economic %>%
      mutate(evtype = factor(evtype)) %>%
      group_by(evtype) %>%
      summarise(health_damage = sum(health_damage, na.rm = TRUE),
             economic_damage = sum(economic_damage, na.rm = TRUE))

graph2 = health_economic_total %>%
      arrange(desc(health_damage)) %>%
      head(20) %>%
      mutate(evtype = fct_reorder(evtype, health_damage)) %>%
      ggplot(aes(evtype, health_damage))+
      geom_col(aes(fill = evtype))+
       coord_flip()+
      labs(title = 'Deadliest events in the US',
           subtitle = 'Tornado events recorded highest injury and fatality count',
           x = '',
           y = 'count of injuries/deaths')+
      scale_y_continuous(labels = scales::comma)+
      theme_minimal()+
      theme(legend.position = 'none')

graph2
```

### Question3: Most economically damaging events

```{r}
top10_property_damage = economic_damages %>%
      mutate(evtype = factor(evtype), 
             type = factor(type)) %>%
      filter(type == 'propdmg') %>%
      group_by(evtype) %>%
      summarise(damage = sum(damage)) %>%
      arrange(desc(damage)) %>%
      head(10)
      

graph3 = ggplot(top10_property_damage, aes(x = reorder(evtype, damage), y = damage/10^6)) + 
      geom_col(aes(fill = evtype))+
      coord_flip()+
      labs(title = 'Top 10 events with the economic damage',
           subtitle = 'Property damage',
           x = '',
           y = '')+
      scale_y_continuous(labels = scales::comma)+
      theme_minimal()+
      theme(legend.position = 'none')

top10_crop_damage = economic_damages %>%
      mutate(evtype = factor(evtype), 
             type = factor(type)) %>%
      filter(type == 'cropdmg') %>%
      group_by(evtype) %>%
      summarise(damage = sum(damage)) %>%
      arrange(desc(damage)) %>%
      head(10)

graph4= ggplot(top10_crop_damage, aes(x = reorder(evtype, damage), y = damage/10^6)) + 
      geom_col(aes(fill = evtype))+
      coord_flip()+
      labs(subtitle = 'Crop damage',
           x = '',
           y = 'cost (M)')+
      scale_y_continuous(labels = scales::comma)+
      theme_minimal()+
      theme(legend.position = 'none')

graph5 = graph3/graph4
graph5
```













