---
title: "Week 4"
author: "Me"
date: "12/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Caching computation

cacher package in R
- add-on package in R
- evaluates code written in files and stores immediate results in a key value database
- R expressions are given SHA-1 values so that changes can be tracked
- 'cacher' package can be built for distribution
- others can clone an analysis and evaluate subsets of code or inspect data objects

using cacher as an author
- parse R source file; create necessary cache directories and subdirectories
- cycle through each expression in the source file
      - if an expression has never been evaluated, evaluate it and store object in cache database
      - if cache results exist, lazy-load results from cache database and move to next expression
      - if an expression does not create any R objects, add expression to the list of expression where evaluation
         is forced

using cache as a reader
```{r}
library(cacher)
clonecache(id = '092d')

showfiles()

source('top20.R')
```


Cloning an analysis
- local directories are created
- source code files and metadata are downloaded
- data objects are not downloaded by default
- references to data objects are loaded and corresponding data can be lazy loaded

Running code
- runcode function executes code in the source file
- by default, expressions that results in an object are not run and the resulting objects is lazy loaded into the workspace.

Checking code
- check code function evaluates all expressions from scratch
- results of evaluation are checked against stored results to see if results are the same as what authors calculated

cacher summary
- authors can create packages from data analyses for distribution
- readers can use cacher package to inspect other data analyses by examining cache computations

# Case study: Air pollution

Identify harmful elements of air pollution.

NMAPPS reproducibility
- data are made available at the internet
- research results and software are also available
- many studies have been conducted based on the public data

