---
title: "Week 2"
author: "Me"
date: "12/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Coding standards in R
1. always use text editor. 
2. indent your code
3. limit width of your code
4. Limit length of functions. 


# Markdown
- simplified markup language. Just focus on content
- text to HTML conversion tool for web writers. Allows you to write using easy to read and write text format to valid XHTML or HTML

Types
*italics*
**bold**
##Headings
- Unordered
- List
- Using hypen

1. Ordered
2. List
3. Using numbers

Links
[Rstudio](http://www.rstudio.com)

Advanced linking
I spend so much time reading [R bloggers][1] and [Simply Statistics][2]!
[1]: http://www.r-bloggers.com/ 'R bloggers'
[2]: http://simplystatistics.org/ 'Simply Statistics'

New lines are created after double space

First Line

Second Line


# R markdown

- integration of R code with markdown
- Allows one to create documents containing 'live' R code
- R code is evaluated is part of markdown
- Results from R code are inserted into markdown document.
- Core tool for literate statistical programming
- can be converted to standard markdown using knitr package
- markdown can be converted to html using markdown package.
- any basic text editor can be used to create a markdown document.
- R markdown -> markdown -> HTML workflow can be easily managed using R studio
- R markdown can be converted to slides using slidify package.


# R markdown demonstration


This is an R markdown file

Here we will load some data

```{r}
library(datasets)
summary(airquality)
```
Let's first make a pairs plot of the data

```{r}
pairs(airquality)
```

Here's a regression model of ozone on wind and temperature

```{r}
fit = lm(Ozone ~ Wind + Solar.R + Temp, data = airquality)
summary(fit)
```

#Knitr

- Authors must undertake considerable effort to put data/results on the web
- readers must download data/results and piece together which data go with which code sections etc

Literate Statistical programming
- article is a stream of text and code
- analysis of code is divided into code of chunks
- presentation code formats results (tables, figures, etc)

How do I make my work reproducible
- decide to do it
- keep track of things, perhaps with a version control
- use software whose operation can be encoded
- dont save output (save raw data and preprocessing code)
- save data in non proprietary formats

Pros of literate programming
- text and code in one place
- data, results automatically update to reflect external changes
- code is live when buildign document

Cons
- can make documents hard to read, especially if there are a lot of codes
- can slow down processing of documents.


# What is knitr
- supports Rmarkdown, latex, HTML as documentation language
- can export PDF, HTML
- built into Rstudio

what is knitr good for?
- manuals
- short/medium length technical documents
- tutorials
- report
- data preprocessing documents/ summaries


Not good for
- long research articles
- complex time consuming computations
- documents that require precise formatting. 


Options
echo = FALSE : means no code echoed
results = 'hide' : hide results


can also change figures 

```{r scatterplot, fig.height=4}
x = rnorm(100)
y = x + rnorm(100, sd = .5)

par(mar = c(5,4,1,1), las = 1)
plot(x,y, main = 'Simulated data')
```

Making tables with xtable
```{r}
library(datasets)
data("airquality")
fit = lm(Ozone ~ Wind + Temp + Solar.R, data = airquality)
```

Here is a table of regression coefficients
```{r}
library(xtable)
xt = xtable(summary(fit))
print(xt)
```

Setting global options
- apply settings to every code chunks

```{r setoptions, echo = FALSE}
opts_chunk$set(echo = FALSE, results = 'hide')
```


Some common options

Output
- results: asis, hide
- echo : TRUE, FALSE

Figures
- fig.height : numeric
- fig.width : numeric

Caching computations
- what if one chunk takes a long time to run?
- all chunks have to be recomputed everytime you re-knit the file
- cache = TRUE option can be set on chunk by chunk basis to store results
- after first run, results are loaded from cache.

Caching caveats
- if the data or code changes, need to rerun cached code chunks
- dependencies are not checked
- chunks with side effects may not be cacheable

































