---
title: "Week 2"
author: "Me"
date: "12/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Coding standards in R
1. always use text editor. 
2. indent your code
3. limit width of your code
4. Limit length of functions. 


# Markdown
- simplified markup language. Just focus on content
- text to HTML conversion tool for web writers. Allows you to write using easy to read and write text format to valid XHTML or HTML

Types
*italics*
**bold**
##Headings
- Unordered
- List
- Using hypen

1. Ordered
2. List
3. Using numbers

Links
[Rstudio](http://www.rstudio.com)

Advanced linking
I spend so much time reading [R bloggers][1] and [Simply Statistics][2]!
[1]: http://www.r-bloggers.com/ 'R bloggers'
[2]: http://simplystatistics.org/ 'Simply Statistics'

New lines are created after double space

First Line

Second Line


# R markdown

- integration of R code with markdown
- Allows one to create documents containing 'live' R code
- R code is evaluated is part of markdown
- Results from R code are inserted into markdown document.
- Core tool for literate statistical programming
- can be converted to standard markdown using knitr package
- markdown can be converted to html using markdown package.
- any basic text editor can be used to create a markdown document.
- R markdown -> markdown -> HTML workflow can be easily managed using R studio
- R markdown can be converted to slides using slidify package.


# R markdown demonstration


This is an R markdown file

Here we will load some data

```{r}
library(datasets)
summary(airquality)
```
Let's first make a pairs plot of the data

```{r}
pairs(airquality)
```

Here's a regression model of ozone on wind and temperature

```{r}
fit = lm(Ozone ~ Wind + Solar.R + Temp, data = airquality)
summary(fit)
```




